---
title: "BRIC_project"
author: 
  name: Manuel Schreiber
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    df_print: paged
  html_notebook: default
  latex_engine: pdflatex
  pdf_document: default
  header-includes:
   - \usepackage{amsmath}
   - \usepackage{amsfonts}
---

**BRIC Project**

```{r}
# clear workspace
##rm(list=ls())
```


getting the current WD

```{r}
getwd()
```

changing the WD

```{r}
##setwd("/Users/Manu/Desktop/TUM_Master_Mgt_Technology/TUM_SS_21/Empirical Asset Pricing seminar/Seminar Thesis/BRIC_data")
```

# Loading Libraries

```{r}
# loading libraries
library(data.table) # extension of the data.frame package. It is widely used for fast aggregation of large datasets, low latency add/update/remove of columns, quicker ordered joins, and a fast file reader.
library(dplyr) # data manipulation package
library(lubridate)
library(zoo) # methods for totally ordered indexed observations. It aims at performing calculations containing irregular time series of numeric vectors, matrices & factors

library(stats)
library(utils)
library(tidyverse)

```


# Loading in the R.data

Data column descriptions (Worldscope):
https://www.professors.wi.tum.de/fileadmin/w00bca/fm/Worldscope_Data_Definition_Guide_Issue_15.pdf

https://docs.google.com/spreadsheets/d/1YtuJiv60Q6nKIaFJLQY60sGQErbsl_8nvPdUHmvO8vM/edit?usp=sharing


```{r}
memory.limit(9999999999)
# loading R.data BRIC monthly
load("C:/Users/johan/Documents/BRIC_Asset_Pricing_Project/BRIC_monthly.RData")

# loading R.data BRIC monthly
load("C:/Users/johan/Documents/BRIC_Asset_Pricing_Project/BRIC_static.RData")

# loading R.data BRIC monthly
load("C:/Users/johan/Documents/BRIC_Asset_Pricing_Project/BRIC_yearly.RData")

```
# Clearing data from dates before July 1994 and NA s in RET.USD

```{r}
# save raw data and adjust others to the required date range
BRIC.monthly.raw <- BRIC.monthly
BRIC.monthly <- subset(BRIC.monthly, Date >= "1994-06-30" & Date <= "2019-01-01")
BRIC.monthly <- na.omit(BRIC.monthly,cols = "RET.USD")

BRIC.yearly.raw <- BRIC.yearly
BRIC.yearly <- subset(BRIC.yearly, YEAR >= "1994" & YEAR <= "2019")


```

ISIN: International Security Identification Number (stock identifier)
ESTAT: active vs inactive company (publicly listed or not)
Id: join column with BRIC.yearly dataframe
INDM: industry sector code

GEOGN: geographic group name
GEOLN: geographic location

List of Database codes:
https://www.bwl.uni-mannheim.de/media/Lehrstuehle/bwl/Maug/Database_info/Datastream_dataypes.pdf

ID: ??
Country: 4 BRIC COUNTRY CODES
ICBSUC: industrial classification benchmark
https://link.springer.com/content/pdf/bbm%3A978-3-8350-9531-1%2F1.pdf

WC07021: SIC(standard industrial classification) primary code from Worldscope
W05651: Common Shares Traded - Annual (Security)


*TO DO's*:

- Calculate Benchmark (market-cap weighted big stocks of the BRIC region)
- Calculate break points on big stocks
- decide on strategy and implement it
- calculate portfolio characteristics
- visualize and showcase strategy performance
- regress strategy PF on FF5FM and momentum for style exposure analysis
- Literature 


*Project Dates*:
Thesis submission: June 21
Final presentation: June 07

*Strategy*:
GDP weighted countries; stock level: max sharpe ratio, min volatility, equal sector weights or quotas, momentum?


EDA

```{r}
# finding the number of NA values by column in the dataframe
BRIC.monthly %>% mutate_all(is.na) %>% summarize_all(sum)

```

To DO: compute correlations by sector (10 sectors)

```{r}
# regular correlation matrix of all (four) numeric attributes
cor(select(BRIC.static, where(is.numeric)))
```


**Some conventions:**

Characteristic should be calculated as in Hanauer & Lauterbach (2019) or in Hanauer (2020)

Big stocks should be defined as the biggest stocks which together account for 90% of a
country's aggregated market capitalization
Benchmark should be defined as the cap-weighted universe of big stocks
Returns should be in USD
Breakpoints (for Fama-French factors) should be calculated on big stocks (as in the
excursus) but both small and big stocks go into the factor calculation.

# Benchmark Calculation

```{r}
length(BRIC.static$Id)

length(BRIC.yearly$Id)

```

```{r}
# joining the monthly and the static dataframe by the Id column (retainig all rows of BRIC.yearly)
df_combined <- left_join(x=BRIC.monthly,y=BRIC.static,by="Id")


head(df_combined)



```

```{r}
summary(df_combined$MV.USD)
```


```{r}
## finding the number of NA values by column in the dataframe
BRIC.yearly %>% mutate_all(is.na) %>% summarize_all(sum)
```



```{r}
# markt cap (USD): WC07210

## finding the number of NA values by column in the dataframe
df_combined %>% mutate_all(is.na) %>% summarize_all(sum)
```




Idea: calculate 90% of the sum of all stock market caps per country as a time series
(store the result in a column )

```{r}
# Big stocks should be defined as the biggest stocks which together account for 90% of # a country's aggregated market capitalization


# grouping the mcap values of individual stocks by country
mcap_country <- df_combined %>%
  filter(df_combined$MV.USD != "NA") %>% 
  group_by(country.x, ym) 

# computing the sum of the individual stock's market caps by country
country_mcap <- mcap_country %>%
group_by(country.x, ym) %>%
    summarize(mc_sum = sum(MV.USD))

country_mcap


```

```{r}
summary(country_mcap$mc_sum[country_mcap$country.x == "BRA"])
summary(country_mcap$mc_sum[country_mcap$country.x == "RUS"])
summary(country_mcap$mc_sum[country_mcap$country.x == "IND"])
summary(country_mcap$mc_sum[country_mcap$country.x == "CHN"])
```

# FF5FM Calulation new

```{r}
## 1 month t-bill rate
# load data sheet from French's website
library(readr)
FFData <- read_csv("FF_Research_Data_5_Factors_2x3.CSV", 
     skip = 2)

# adding a ym column to risk free rate data
one_m_tbill <- as.data.frame(FFData[c("X1","RF")][1:693,])

one_m_tbill$ym<-as.yearmon(one_m_tbill$X1, "%Y %m")

# merge risk-free rate (1 month treasury bill rate) with monthly data
BRIC.monthly_full <- left_join(x = BRIC.monthly, y = one_m_tbill, by = "ym")

# make rf column numeric
BRIC.monthly_full$RF <- as.numeric(BRIC.monthly_full$RF)

```

# calculate RMRF

```{r}
BRIC.monthly_full$RMRF <- BRIC.monthly_full$RET.USD - BRIC.monthly_full$RF
```

# Calculate SMB and HML
comment: na values appear for size calculation, which differs from what in the excursus
note: we omitted the na's for further calculation --> has to be clarified in the Q&A session with Hanauer

we didn't calculate the lagged MV yet lines 271,315

```{r}
## Determine portfolio breakpoints for Size
# Fama-French take the MV from end-of-June and rebalance yearly
BRIC.monthly_full[,month := month(Date)]
BRIC.monthly_full[,year := year(Date)]
BRIC.monthly_full[,hcjun := ifelse(month>=7,year,year-1)]

# determine size portfolio allocation from July on using data that's public from end-of-June on

# we didn't calculate the lagged MV yet 

setorder(BRIC.monthly_full,Date,-MV.USD)
hlpvariable <-  BRIC.monthly_full[month==7 & !is.na(MV.USD),
                .(Size = ifelse((cumsum(MV.USD)/sum(MV.USD))>=0.9,"Small","Big"),Id),
                               by=year]

# Merge the size portfolio allocation back from July Y to June Y+1
panel_country <- merge(BRIC.monthly_full,hlpvariable,
                       by.x=c("hcjun","Id"),
                       by.y=c("year","Id"),
                       all.x=T)

panel_country_noNA <- na.omit(panel_country,cols = "Size") 

# merge yearly BM data with our table
BRIC.yearly_helper <- subset( BRIC.yearly, select = c("Id","country","ICBSUC","YEAR","WC03501","WC07210","WC01001","WC01051","WC01101","WC01251","WC02999") )
BRIC.yearly_helper$BM <- BRIC.yearly_helper$WC03501 / BRIC.yearly_helper$WC07210

# delete NA s from BM column
BRIC.yearly_helper <- na.omit(BRIC.yearly_helper,cols = "BM")

summary(BRIC.yearly_helper$BM)

# eliminating unused columns
panel_country_noNA <- subset(panel_country_noNA, select = c("hcjun","Id","country","MV.USD","RET.USD","ym","RF","month","year","Size","Date"))

panel_country_BM <- inner_join(x = panel_country_noNA,y = BRIC.yearly_helper, by = "Id")


# Determine the B/M breakpoints based on big stocks only
hlpvariable2 <- panel_country_BM[month==7 & !is.na(BM) & Size=="Big", .(bm_bb30 = quantile(BM , probs = c(0.3), na.rm=T),
                                                                        bm_bb70 = quantile(BM , probs = c(0.7), na.rm=T)),by=year]
              
# Merge the B/M portfolio allocation back from July Y to June Y+1
panel_country_new <- merge(panel_country_BM,hlpvariable2,
                       by.x=c("hcjun"),
                       by.y=c("year"),
                       all.x=T)

panel_country_new[ , pf.bm := ifelse(BM>bm_bb70,"High",ifelse((BM<=bm_bb70 & BM>bm_bb30),"Neutral",ifelse(BM<=bm_bb30,"Low",NA)))]

panel_country_new[, SIZE_VALUE := paste0(Size,".",pf.bm)]

# we used MV.USD instead of the lagged MV

portfolio_returns <- panel_country_new[!is.na(Size) & !is.na(pf.bm)] %>% # this operator nests functions
  group_by(Date,SIZE_VALUE) %>% # do "everything" for the groups specified here
  summarize(ret.port = weighted.mean(RET.USD,
                                     MV.USD)) %>% # vw returns using lagged mcap
  spread(SIZE_VALUE,ret.port) %>% # create one column for each group
  mutate(
    Small = (Small.High + Small.Neutral + Small.Low)/3, # just exemplary
    Big = (Big.High + Big.Neutral + Big.Low)/3,
    SMB = Small-Big,
    High = (Small.High + Big.High)/2,
    Low = (Small.Low + Big.Low)/2,
    HML = High-Low
  )

portfolio_returns <- as.data.table(portfolio_returns)

head(portfolio_returns)
 
```
 
# RMW
profitability: OP/BE: Operating profits-to-book equity. We measure operating profits-to-book equity as in Fama and French (2015) as sales or revenues (WC01001) minus cost of goods sold (WC01051), minus selling, general, and administrative expenses (WC01101), minus interest expense (WC01251); all divided by book equity. (Hanauer, 2019, p. 284)
```{r}
# use BRIC.yearly_helper



```





# CMA
investment: As in Cooper et al. (2008), we measure asset growth in June of year y as the percentage change in total assets (WC02999) from fiscal year ending in calendar year y−2 to fiscal year ending in calendar year y−1.
```{r}
# use BRIC.yearly_helper



```

# Strategy


```{r}
#install.packages("plm")
library(plm)

strategy_df <- pdata.frame(BRIC.monthly,index = c("Id","ym"))
#strategy_df <- transform(strategy_df,l_value=lag(value,1))

#for(i in 1:length(strategy_df$Id)){
#  sd(strategy_df$RET.USD)
#strategy_df$std_value <- 
#}

sd.y <- as.numeric()
for(i in 1:10){
  sd.y[i] <- sd(strategy_df$RET.USD[i])
}
sd.y
# [1] 0.9681038 0.3783425 1.1031686 1.1799477 0.6867556 1.6987277 
# [7] 1.8859794 1.4993717 1.2956209 1.1116502

ag <- aggregate(. ~ Id, strategy_df, function(x) c(mean = mean(x), sd = sd(x)))

# jjs

```



```{r}
# sorting the stocks by market cap and extracting the ISINs of the biggest ones that make up 90% of a country's market cap

# RUSSIA
# 90% of the market cap in Russia: 
0.9*country_mcap$mc_sum[4]

# subsetting all Russian stocks
russian_stocks <- as.data.frame(mcap_country) %>%
   filter(country.x == "RUS") 

# sorting the dataframe of russian stocks by decreasing market cap
russian_stocks <- russian_stocks[order(russian_stocks$WC07210,decreasing = TRUE),]
 
# filtering out the big stocks which sum up to 90% of the market cap of Russian stocks
big_stocks_RUS <- russian_stocks[cumsum(russian_stocks$WC07210) <= 0.9*country_mcap$mc_sum[4],]

big_stocks_RUS



```


idea:

big_stocks <- mcap_country %>%
  select(ISIN) %>%
 sort(mcap_country$WC07210, decreasing=T)  %>%
  filter(mcap_country$WC07210 <= 0.9*country_mcap$mc_sum)

big_stocks

```{r}
# sorting the stocks by market cap and extracting the ISINs of the biggest ones that make up 90% of a country's market cap

# RUSSIA
# 90% of the market cap in Russia: 
0.9*country_mcap$mc_sum[4]

# subsetting all Russian stocks
russian_stocks <- as.data.frame(mcap_country) %>%
   filter(country.x == "RUS") 

# sorting the dataframe of russian stocks by decreasing market cap
russian_stocks <- russian_stocks[order(russian_stocks$WC07210,decreasing = TRUE),]
 
# filtering out the big stocks which sum up to 90% of the market cap of Russian stocks
big_stocks_RUS <- russian_stocks[cumsum(russian_stocks$WC07210) <= 0.9*country_mcap$mc_sum[4],]

big_stocks_RUS



  
```


# Calculating the size portfolios




